---
title: "Penguins ANOVA"
author: "Anna Calle"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read-in csv data

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")


```


Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring
Note: already in tidy format: one variable per column and one observation per row
New functions: with +table for easy counting over multiple factors

Update species name using case_when

```{r}
counts <- with(penguins, table(species, sex))
penguins2 <- penguins %>%
  mutate(common_name =
           case_when(
             species ==" Chinstrap penguin (Pygoscelis antarctica)"~"Chinstrap",
         species == "Gentoo penguin (Pygoscelis papua)"~"Gentoo",
         species == "Adelie Penguin (Pygoscelis adeliae)"~"Adelie")
  ) %>% 
  select(-species) %>% 
  filter(sex=="FEMALE")

```

-Exploratory graphs
-Summary statistics
-Levenes's test fro equal variances
-one-way ANOVA
-post-hoc testing

```{r}
hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~common_name)
hists

qqplot <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +
  facet_wrap(~common_name)
qqplot

# Levene's test for equal variances ( >2 groups...if you have two groups , use the F-test for equal variances)

#We'll use the function leveneTest from the 'car' package

# H0: Variances are equal
# HA: Variances are not equal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# We retain the null hypothesis of equal variances (p = 0.06)

variance_table <- penguins2 %>% 
  group_by(common_name) %>% 
  summarise(
    variance = var(flipper_length)
  )

variance_table

```








